Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2012-10-23 documents/mixtures/mixture_models.tex
Address "DSTN" comments, fix affiliation, mention Bosch's work, cite and thank him, mention tractor example figs in the text

% This file is part of the Tractor project.
% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton).
% All rights reserved.

% to-do
% -----
% - DSTN: check sentences I say about ``ten times faster''.
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - DSTN: make figures showing use of the models in a real situation; see last figure
% - HOGG: does the Lupton paper give the profiles, or should we cite the code directly?
% - HOGG: need Bundy reference
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - HOGG: send to Lupton and Bundy for comments
% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
% - DSTN: is there anyone we need to ack on your side, or granting agencies?
% - HOGG: should we make Machine Readable Tables for PASP?

@@

\title{Replacing standard galaxy profiles with \\ mixtures of Gaussians}
\author{David W. Hogg\altaffilmark{1,2,3} \&
        Dustin Lang\altaffilmark{4,5}}
\altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Phyiscs, New York University, 4 Washington Place, New York, NY 10003, USA}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{4}{Princeton University Observatory, Princeton, NJ 08544, USA}
\altaffiltext{5}{Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}

@@

We need good performance in two-dimensional image synthesis (for
fitting) because, with \thetractor\ (Lang \etal, forthcoming), we are
building a comprehensive model of all the imaging data we have; this
will have models of many millions of galaxies in imaging that contains
on the order of $10^{13}$ pixels.  We are basing our models on the
\project{Sloan Digital Sky Survey} (\sdss) Catalog galaxy models,
which only include exponential, de~Vaucouleurs, and composite (mixture
of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
models are small modifications of these (\citealt{lupton}); details
below.  So our goal here is to provide replacements for these models
in order to improve the performance of galaxy image modeling and
analysis software.  Some of the models we present have been used
previously in tools for precise photometry (\citealt{bundy}); they
could also be used to speed other image-fitting systems, like the very
successful \project{GALFIT} (\citealt{galfit}).

@@

DSTN... In \figurename~\ref{fig:example} we show...

@@

\acknowledgements It is a pleasure to thank
      Brendon Brewer (Auckland)
  and Kevin Bundy (IPMU)
for valuable comments.  This work was supported in part by NASA (grant
NNX12AI50G) and the NSF (grant IIS-1124794).  All the results shown in
the figures and tables and more are available in machine-readable form
from author DWH upon request.

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K.~??
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\end{thebibliography}

@@

\clearpage
\begin{figure}
\setlength{\figurewidth}{0.33\textwidth}
% DSTN:  This should be a 3-across by 2-down figure,
\begin{tabular}{@{}c@{}c@{}}
\includegraphics[width=\figurewidth]{trdemo-01} &
\includegraphics[width=\figurewidth]{trdemo-02} \\
\includegraphics[width=\figurewidth]{trdemo-03} &
\includegraphics[width=\figurewidth]{trdemo-06} \\
\includegraphics[width=\figurewidth]{trdemo-08} &
\includegraphics[width=\figurewidth]{trdemo-09}
\end{tabular}
\caption{Demonstration of use of the profiles, or the implicit
  generative model in this \documentname.  \textsl{top-left:} The
  circular dimensionless $M^{\luv}=8$ mixture-of-Gaussian
  approximation to the luv profile, represented on a very fine pixel
  grid. \textsl{top-right:} The ellipse representing the non-trivial
  affine transformation to be applied to the circular, dimensionless
  profile.  \textsl{middle-left:} The sheared profile.
  \textsl{middle-right:} A $K=3$ mixture-of-Gaussian model of the
  pixel-convolved point-spread function, represented on the very fine
  pixel grid.  \textsl{bottom-left:} The sheared profile convolved
  with the PSF, represented on the very fine pixel grid.
  \textsl{bottom-right:} The sheared luv convolved with the PSF, but
  now shown on a realistic pixel grid.  Because by assumption the PSF
  is a pixel-convolved PSF, the representation on the realistic grid
  is found simply by interpolating to the pixel centers the convolved
  high-resolution model.\label{fig:example}}
\end{figure}

>>>>

% This file is part of the Tractor project.
% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton->CMU).
% All rights reserved.

% to-do
% -----
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - HOGG: need Bundy reference
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - HOGG: send to Lupton and Bundy for comments
% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
% - HOGG: Cite paper mentioned by Bosch?
% - HOGG: should we make Machine Readable Tables for PASP?

@@

\title{Replacing standard galaxy profiles with \\ mixtures of Gaussians}
\author{David W. Hogg\altaffilmark{1,2,3} \&
        Dustin Lang\altaffilmark{4,5}}
\altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{4}{Princeton University Observatory, Princeton, NJ 08544, USA}
\altaffiltext{5}{McWilliams Center for Cosmology, Carnegie Mellon University, Pittsburgh, PA 15213, USA}

@@

We need good performance in two-dimensional image synthesis (for
fitting) because, with \thetractor\ (Lang \etal, forthcoming), we are
building a comprehensive model of all the imaging data we have; this
will have models of many millions of galaxies in imaging that contains
on the order of $10^{13}$ pixels.  We are basing our models on the
\project{Sloan Digital Sky Survey} (\sdss) Catalog galaxy models,
which only include exponential, de~Vaucouleurs, and composite (mixture
of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
models are small modifications of these (\citealt{lupton}); details
below.  So our goal here is to provide replacements for these models
in order to improve the performance of galaxy image modeling and
analysis software.  Some of the models we present have been used
previously in tools for precise photometry (\citealt{bundy}), and are
being used in the Large Synoptic Survey Telescope (LSST) analysis
codes along with shapelet representations for the PSF
(\citealt{bosch}).  These models could also be used to speed other
image-fitting systems, like the very successful \project{GALFIT}
(\citealt{galfit}).

@@

In \figurename~\ref{fig:example} we show how we are using these MoG
approximations in \project{The Tractor}---a generative modeling
framework for measuring astronomical objects---to render PSF-convolved
galaxy images.

@@

\acknowledgements It is a pleasure to thank
      Brendon Brewer (Auckland),
      Jim Bosch (Princeton),
  and Kevin Bundy (IPMU)
for valuable comments.  This work was supported in part by NASA (grant
NNX12AI50G) and the NSF (grant IIS-1124794).  All the results shown in
the figures and tables and more are available in machine-readable form
from author DWH upon request.

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bosch(2010)]{bosch}
Bosch,~J., 2010, \aj, 140, 870
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K.~??
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\end{thebibliography}

@@

\clearpage
\begin{figure}
\setlength{\figurewidth}{0.33\textwidth}
\begin{tabular}{@{}c@{}c@{}}
\includegraphics[width=\figurewidth]{trdemo-01} &
\includegraphics[width=\figurewidth]{trdemo-02} \\
\includegraphics[width=\figurewidth]{trdemo-03} &
\includegraphics[width=\figurewidth]{trdemo-06} \\
\includegraphics[width=\figurewidth]{trdemo-08} &
\includegraphics[width=\figurewidth]{trdemo-09}
\end{tabular}
\caption{Demonstration of use of the profiles, or the implicit
  generative model in this \documentname.  \textsl{top-left:} The
  circular dimensionless $M^{\luv}=8$ mixture-of-Gaussian
  approximation to the luv profile, represented on a very fine pixel
  grid. \textsl{top-right:} The ellipse representing the non-trivial
  affine transformation to be applied to the circular, dimensionless
  profile.  \textsl{middle-left:} The sheared profile.
  \textsl{middle-right:} A $K=3$ mixture-of-Gaussian model of the
  pixel-convolved point-spread function, represented on the very fine
  pixel grid.  \textsl{bottom-left:} The sheared profile convolved
  with the PSF, represented on the very fine pixel grid.
  \textsl{bottom-right:} The sheared luv convolved with the PSF, but
  now shown on a realistic pixel grid.  Because by assumption the PSF
  is a pixel-convolved PSF, the representation on the realistic grid
  is found simply by interpolating to the pixel centers the convolved
  high-resolution model.\label{fig:example}}
\end{figure}
